How fluid-mechanical erosion creates anisotropic porous media

### 2022-12-15 docs/PhysicsPaper/Referee Reports/Ref 2 Response/Ref2.tex
NM

\noindent
Thank you for bringing up this point. We have no explicitly stated that the computations are two-dimensional in the abstract, the introduction, and in a few other places in the paper. We also included a brief discussion at the beginning of Section 3 on modifications that would be necessary for 3D. In short, many aspects of the numerical methods (not only the quadrature method) would need to change substantially.

>>>>

\noindent
Thank you for bringing up this point. We have no explicitly stated that the computations are two-dimensional in the abstract, the introduction, and in a few other places in the paper. We also included a brief discussion at the beginning of Section 3 on modifications that would be necessary for 3D. In short, many aspects of the numerical methods (not only the quadrature method) would need to change substantially.
\\ \\

\noindent \comment{Reviewer \#2: 
- In the introduction, first paragraph, it is stated: ``Typical materials are more permeable to flow in the longitudinal (or horizontal) direction than transversely…" - what does this even mean? What determines the longitudinal direction in such a typical material? The statement (and the short discussion that follows) seems almost circular to me - the material is more permeable in the direction in which it admits flow? That's almost the definition of permeability.
\\ }

\noindent
It is the direction of the pressure gradient that primarily determines the direction of flow. The direction of the pressure gradient could change with time, and so, in principle, a porous material with fixed microstructure (and hence fixed permeability tensor) could admit flow in any direction. For example, a porous material might be isotropic, meaning that it admits flow in the vertical direction just as easily as it does in the horizontal direction, or any direction for that matter. The actual flow direction at any given time would thus be determined entirely by the pressure gradient that is imposed. As a second example, a porous medium might be composed of slender and primarily vertically-aligned particles. This material would preferably admit flow in the vertical direction over the horizontal direction, but if the pressure gradient is horizontal, the flow will be horizontal. Hence, the direction of the anisotropy in permeability does not determine the flow direction, although it is an influencing factor.

We have slightly rephrased this statement, removing the terms longitudinal and transversely and replacing them with horizontal and vertical, to prevent confusion.
\\ \\

\noindent \comment{Reviewer \#2: 
- Section 2, page 2, below equations (1): "To nondimensionalize, we set the fluid viscosity to unity, $\mu=1$." This is a lazy way to nondimensionalize. The authors have already gone half-way to nondimensionalizing by this stage, without saying so, by choice of the computational domain. Why not just do it properly and say what the appropriate scalings for all physical quantities are? Then $\mu$ won't appear in any of the equations (which currently have an uncomfortable hybrid feel - are they meant to be dimensionless, or not?).
\\ }

\noindent
We have taken the referee's advice by removing $\mu$ from the governing equations and stating the scales used in the non-dimensionalization. Please see the revised Section 2. We note that we have kept $\mu$ in Eq.~(9), and shortly thereafter set it to unity, because it may be useful for future work to see where the viscosity enters Darcy's law.
\\ \\

\noindent \comment{Reviewer \#2: 
- Section 3.2, first line: "…vorticity, which reduces to shear stress $\tau$ on solid boundaries" - is that right?
\\ }

\noindent
Yes, the vorticity reduces to the shear stress, up to a sign, on a no-slip boundary. We note that the no-slip part is essential for this to be true. It can be seen in 2D by, without loss of generality, taking a horizontal surface and evaluating the vorticity $\omega =  v_x - u_y$ on the no-slip surface. By the no-slip assumption, $v_x = 0$, and so $\omega = -u_y$ which is the same, up to a sign, as the shear stress $\tau$. For any other alignment of the surface, the coordinate system can simply be rotated. Since we use $\abs{\tau}$ in the erosion law, the sign makes no difference.
\\ \\

\noindent \comment{Reviewer \#2: 
- Section 4.1, diagonal permeability tensor. I understand why the authors take the approach that they do, and most likely it's justified. However, the authors don't ever check the consistency of the diagonal assumption. How would they go about checking this? Could this be commented on?
\\ }

\noindent
This is a great question and one that we have thought about. Unfortunately, it would be difficult to check the diagonal assumption in the current computational setup since rotating the configuration of bodies by any angle that is not a multiple of $90^\circ$ does not guarantee the bodies would remain in the square domain. Imagining a nice computational domain that is not subject to this limitation is tricky. In principle, we could confine the bodies to a circular sub-domain (as opposed to a square sub-domain) with the whole domain, including the buffer region, remaining a rectangle. This setup has its own pitfalls though. For one, the flow would impinge on the bodies in a non-uniform way. For example, the bodies near the left-most and right-most extreme of the circular subdomain would experience the `leading front' of the flow profile.
\\ \\

\noindent \comment{Reviewer \#2: 
- Paragraph below eq.(36) - "keep the outer geometry fixed and rotate the inner configuration of bodies" - in the model introduction, the authors described the "buffer zones" [-3,-1], [1,3] in the x-direction, used to allow the flow to adjust to the presence of the bodies (and presumably give more reliable indicators for the anisotropy measures calculated). But the computational domain in the y-direction is only given as [-1,1]. Do the authors extend it for the y-permeability calculation, to get similarly reliable estimates? Or maybe I'm misunderstanding the procedure used.
\\ }

### 2022-12-15 docs/PhysicsPaper/StokesErosion.tex
NM

Flow-induced erosion acts across a range of scales in the natural world, from massive geological structures sculpted by wind or water~\cite{abrams2009growth, perkins2015amplification, mac2020ultra, sharma2022alcove, mac2022morphological}, to mesoscopic patterns formed by surface or internal flows~\cite{berhanu2012shape, bertagni2021hydrodynamic, weady2022anomalous}, and down to granular and porous networks slowly disintegrating in groundwater flows~\cite{chiu2020viscous, szymczak2009wormhole, jager2017channelization, grodzki2019reactive, bizmark2020multiscale, derr2020flow, zareei2022temporal}. The associated nonlinear feedback between changing shapes and the surrounding flows can imprint across all of these scales, affecting large-scale features as well as small-scale ones, such as the microstructure of porous materials. Porous media encountered in nature typically exhibit material {\em anisotropy} in that they permit seepage flow in certain directions more easily than in others. Typical materials are more permeable to flow in the longitudinal (or horizontal) direction than transversely by a factor of 5--20~\cite{bear1988dynamics, anderson2015applied}. Most commonly, this material anisotropy is attributed to the sedimentation process, in which, due to the fluid-structure interaction, non-spherical particles tend to settle with their long axis parallel to the plane of deposition~\cite{bear1988dynamics}. Controlled experiments, however, have not been performed to test this hypothesis, and other mechanisms may be at work. Here, we use numerical simulations to examine an alternative, and possibly complementary, mechanism: namely, that the flow-induced erosion of the medium's solid constituents contributes to its overall anisotropy.

>>>>

Flow-induced erosion acts across a range of scales in the natural world, from massive geological structures sculpted by wind or water~\cite{abrams2009growth, perkins2015amplification, mac2020ultra, sharma2022alcove, mac2022morphological}, to mesoscopic patterns formed by surface or internal flows~\cite{berhanu2012shape, bertagni2021hydrodynamic, weady2022anomalous}, and down to granular and porous networks slowly disintegrating in groundwater flows~\cite{chiu2020viscous, szymczak2009wormhole, jager2017channelization, grodzki2019reactive, bizmark2020multiscale, derr2020flow, zareei2022temporal}. The associated nonlinear feedback between changing shapes and the surrounding flows can imprint across all of these scales, affecting large-scale features as well as small-scale ones, such as the microstructure of porous materials. Porous media encountered in nature typically exhibit material {\em anisotropy} in that they permit seepage flow in certain directions more easily than in others. Typical materials are more permeable to flow in the \edit{horizontal} direction than \edit{the vertical} by a factor of 5--20~\cite{bear1988dynamics, anderson2015applied}. Most commonly, this material anisotropy is attributed to the sedimentation process, in which, due to the fluid-structure interaction, non-spherical particles tend to settle with their long axis parallel to the plane of deposition~\cite{bear1988dynamics}. Controlled experiments, however, have not been performed to test this hypothesis, and other mechanisms may be at work. Here, we use numerical simulations to examine an alternative, and possibly complementary, mechanism: namely, that the flow-induced erosion of the medium's solid constituents contributes to its overall anisotropy.
